---
title: "Comparisons"
editor_options: 
  chunk_output_type: console
---

```{r include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
library(chorddiag)
library(htmlwidgets)
library(igraph)
library(readr)
library(tidygraph)
library(tidyverse)

```

## Comparing Past vs. Present Conditions

Using LANDFIRE’s BpS products, we explore two different ways to visualize past vs. current vegetation patterns.

-   First, we present **changes in broad ecosystem types** using an interactive comparison diagram. To present these broad ecosystem trends more effectively, we combined classifications of Existing Vegetation Types (EVT) into broader categories. 

-   Second, we compare **amounts of succession classes** (past and present) for the most prevalent ecosystems.


 

These differences may or may not be ‘real’ on the ground and/or may not be a management priority.

*Note: number presented when hovering equals acres.*

**PAST** &emsp; &emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp; &emsp; &emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp; &emsp; &emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp; &emsp;&emsp; &emsp; &emsp;&emsp; &emsp;**CURRENT**

```{r chord, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# read in data
  chord_df<- read_csv("data/BPS2EVT_chord.csv")
#view(histFireGVchord)

#convert to matrix
matrix_df <-as.matrix(as_adjacency_matrix(as_tbl_graph(chord_df),attr = "ACRES"))

#clean up matrix (could be cleaner!)
matrix_df = subset(matrix_df, select = -c(1:7))

matrix_df <- matrix_df[-c(8:11),]

#make a custom color pallet #eb4034 (redish) #b0af9e(grey)

# ORIGINAL
groupColors <- c("#f5e942", # current ag
                 "#1d4220", # current coni
                 "#397d3f", # current con-hard
                 "#b0af9e", # current deve
                 "#eb4034", # current exotics 
            
                 "#56bf5f", # current hard
                 "#7db7c7", # current riparian
                 "#56bf5f", # past hard
                 "#397d3f", # past con-hard
                 "#7db7c7" # past riparian

)



#make chord diagram
chord<-chorddiag(data = matrix_df,
                 type = "bipartite",
                 groupColors = groupColors,
                 groupnamePadding = 10,
                 groupPadding = 3,
                 groupnameFontsize = 11 ,
                 showTicks = FALSE,
                 margin=150,
                 tooltipGroupConnector = "    &#x25B6;    ",
                 chordedgeColor = "#363533"
)
chord 

#save then print to have white background
htmlwidgets::saveWidget(chord,
                        "chord.html",
                        background = "white",
                        selfcontained = TRUE
                        )
```



<iframe src="chord.html" height="720" width="720" style="border: 1px solid #464646;" allowfullscreen="" allow="autoplay" data-external=".5"></iframe>

<br>

## Succession classes for most dominant Biophysical Settings




```{r scls chart, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6}

BPS_SCLS2 <- read.csv("data/bpsScls2.csv")

bps_scls_3 <- BPS_SCLS2 %>%
  group_by(Var1) %>%
  mutate(total.count = sum(Freq)) %>%
  ungroup() %>%
  dplyr::filter(dense_rank(desc(total.count)) < 4) %>%
  dplyr::select(c("bps_name", "ref_label",  "currentPercent", "ref_percent")) %>%
  pivot_longer(
    cols = c(`ref_percent`, `currentPercent`), 
    names_to = "refCur", 
    values_to = "Percent"
    )

# order classes
bps_scls_3$ref_label <- factor(bps_scls_3$ref_label, levels= c(
  "Developed",
  "Agriculture",
  "UE",
  "UN",
  "E",
  "D",
  "C",
  "B",
  "A"))




sclasplot <-
  ggplot(bps_scls_3, aes(fill = factor(refCur), y = Percent, x = ref_label)) + 
  geom_col(width = 0.8, position = position_dodge()) +
  coord_flip() +
  facet_grid(. ~BpS) +
  scale_x_discrete(limits = (levels(bps_scls_3$ref_label))) +
  labs(
    title = "Succession Classes past and present",
    subtitle = "3 BpSs selected for illustration. Not all succession classes present in all BpSs",
    caption = "\nData from landfire.gov.",
    x = "",
    y = "Percent")+
  theme_minimal(base_size = 14)+
  theme(plot.caption = element_text(hjust = 0, face= "italic"), #Default is hjust=1
        plot.title.position = "plot", #NEW parameter. Apply for subtitle too.
        plot.caption.position =  "plot") +
  scale_fill_manual(values = c("#3d4740", "#32a852" ), # present (grey), historical (green)
                    name = " ", 
                    labels = c("Present",
                               "Past")) +
  facet_wrap(~bps_name, nrow(3),labeller = labeller(bps_name = label_wrap_gen(width = 15))) +
    theme(panel.spacing = unit(.05, "lines"),
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        strip.background = element_rect(color = "black", size = 1))

sclasplot

```


