---
title: "Fire Dependent Ecosystems on the Northern Shawangunk Ridge"
author: 
 " **Diego Schillaci** - Paul Smith's College,  
  **Randy Swaty** - The Nature Conservancy  
  with substantial code contributions from Myles Walimaa"
---

<font size="4"> 
**These LANDFIRE-Powered Landscape Assessments allow users to start exploring trends, landscape conversion and natural resource management options. **

</font> 


## Goals of this report

[LANDFIRE](https://landfire.gov/){target="blank"} products will be used to to:

1. Characterize and map past and present ecosystem conditions.
2. Explore historical disturbance patterns.
3. Summarize patterns of change and management opportunities.


The following charts, maps, and graphs are based on current LANDFIRE products, and provide a starting point for further analysis.


This report will:

* Provide context for the past and present ecosystem conditions of the Northern Shawangunks.
* Demonstrate the power of data visualization to explore ecological patterns and functions using LANDFIRE products.
* Facilitate an understanding of historical and current conditions on a regional scale.



## Introduction to the Northern Shawangunks


Minnewaska and Sam's Point State Park Preserves, located on the Northern Shawangunk Ridge in the Mid-Hudson Valley of New York State are considered ecologically unique and sensitive areas. The combined ~29,000 acre State Park Preserves represent two of the five land management agencies located on the Northern Shawangunk Ridge. Underlying these parks, a unique geologic composition and formational history has created a mosaic of globally unique ecological communities. Many of the ecological communities on this ridge are fire dependent which brings into focus the necessity for proper fire management techniques to allow these ecosystems to thrive.

**Location of this assessment **
<br>

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}

library(foreign)
library(sf)
library(terra)
library(tidyverse)
library(raster)
library(ggspatial)
library(leaflet)


```


```{r read shapefile, message=FALSE, warning=FALSE, include=FALSE}
#  read shape
shapefile_path <- "/Users/Diego/Leaflet_test/data/minnewaska_state_park.shp"
shape_data <- st_read( shapefile_path )

shape_data <- st_transform(shape_data, crs = 4326)
```

```{r locator map, echo=FALSE, message=FALSE, warning=FALSE}
leaflet(shape_data) %>%
  addTiles() %>% 
  addPolygons(color = "orange", weight = 3, opacity = 0.8, fillOpacity = 0.2) %>%
 
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  addProviderTiles(providers$Esri.WorldTopoMap, group = "Terrain") %>%
  addLayersControl(
    baseGroups = c("Terrain", "Satellite"),
    options = layersControlOptions(collapsed = FALSE)
  )


```


## Land Acknowledgement
The Northern Shawangunks are recognized as the lands of the Munsee Lenape, Mohican and Schaghticoke Peoples.




## Identifying Barriers to Prescribed Fire on the Northern Shawangunk Ridge

The goal of this  project is to produce a report for land managers to support the use of active fire management and suggest methods for improving fire management strategies by:

	Creating updated ecosystem maps using LANDFIRE data to assess ecosystem conditions Pre-European colonization, Current ecosystem types, Current and historical fire regimes and tracking changes in fire adapted communities 

	Interviewing the involved land managers and ecologists to assess where the current management strategies need to be improved for the future

	Create an extension-style document, a web report and a series of informative pamphlets that can be used by land managers and distributed as an educational tool


## How to use this web report

* All maps, input datasets and further assistance can be obtained by contacting the author, [Diego Schillaci](mailto:dschillaci@paulsmiths.edu){target="blank"}.
* Review is ongoing.
* To share this web app, simply share the hyperlink.
* Toggle between dark and light display of this site in the upper right corner of the page.
* Learn more about LANDFIRE and The Nature Conservancy [here](about.qmd){target="blank}. 
