---
title: "Northern Shawangunk Ridge, New York"
author: "Diego Schillaci- Paul Smith's College with substantial code contributions from Randy Swaty & Myles Walimaa"
---

<font size="4"> 
**These LANDFIRE-Powered Landscape Assessments allow users to start exploring trends, landscape conversion and natural resource management options. **

</font> 


## Goals of this demonstration

We will use [LANDFIRE](https://landfire.gov/){target="blank"} products to:

1. Characterize and map past and present ecosystem conditions.
2. Explore historical disturbance patterns.
3. Summarize patterns of change and management opportunities.


The following charts, maps, and graphs are based on current LANDFIRE products, and provide a starting point for further analysis.


This demonstration will:

* Provide context for the past and present ecosystem conditions of the Northern Shawangunks.
* Demonstrate the power of data visualization to explore ecological patterns and functions using LANDFIRE products.
* Facilitate an understanding of historical and current conditions on a regional scale.



## Location of this assessment



<br>

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}

library(foreign)
library(sf)
library(terra)
library(tidyverse)
library(raster)
library(ggspatial)


```


```{r read shapefile, message=FALSE, warning=FALSE, include=FALSE}
#  read shape
shp <- st_read("data/minnewaska_state_park.shp") %>% 
  st_transform(crs = 5070) %>%
  st_union() %>%
  st_sf()
```

```{r locator map, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  annotation_map_tile(type = "osm", zoom = 14) +  # OpenStreetMap basemap
  geom_sf(data = shp, fill = "lightgreen", color = "darkgreen", alpha = 0.6) +
  labs(title = "Minnewaska State Park") +
  theme_minimal() + 
  theme(axis.text = element_blank(),  # Remove lat/lon numbers
        axis.ticks = element_blank())  # Remove axis ticks




```

## How to use this web report

* All maps, input datasets and further assistance can be obtained by contacting the author, [Diego Schillaci](mailto:dschillaci@paulsmiths.edu){target="blank"}.
* Review is ongoing.
* To share this web app, simply share the hyperlink.
* Toggle between dark and light display of this site in the upper right corner of the page.
* Learn more about LANDFIRE and The Nature Conservancy [here](about.qmd){target="blank}. 
